Unifying quasars with clumpy wind models

### 2015-08-27 paper.tex
version for sending to collabs

\begin{abstract} 
% Broad absorption lines (BALs) in the ultraviolet 
% are seen in $\sim20\%$ of quasi-stellar objects (QSOs). 
% Blue-shifted broad absorption lines (BALs) are the most direct evidence of 
% accretion disc `winds' in such systems; mass loaded outflows
% emanating from the disc that may be driven by line forces or
% magnetic processes. 
Various unification schemes for
quasars and luminous active galactic nuclei (AGN) have proposed
that the broad emission line region is roughly cospatial
with broad absorption line (BAL) gas and much of the phenomenology of luminous AGN
can be explained by a simple geometrical picture involving an accretion
disc and associated outflow. Here, we test this paradigm by 
utilising our state-of-the-art radiative transfer code to produce synthetic spectra
from simple biconical disc wind models. In particular, we expand on our previous 
work in which a benchmark model for BAL quasars was produced. 
We have conducted a limited parameter search with the aim of unifying quasar phenomenology,
and arrive at an improved model. The grid is now publicly available.
We find that a simple treatment of clumping (`microclumping') 
allows for a more realistic X-ray luminosity in the model by lowering the 
ionization parameter. We examine the X-ray properties of this new model
and find good agreement with existing X-ray samples of AGN and QSOs.
We find that the dense, X-ray heated wind 
produces strong H recombination and collisionally excited resonance 
line emission to emerge at the low inclination angles, 
which represent quasars within this unification scenario.
% We also treat Hydrogen using a `macro-atom' approach in order to 
% examine the effect of recombination on Hydrogen emission lines, and
% this results in significant line emission in Lyman~$\alpha$ and the Balmer series.
However, we are unable to reproduce the remarkably similar of line-to-continuum
ratios between BAL and non-BAL quasar composites.
This is due to a fundamental constraint arising
from the anisotropy of emission from a classical thin disc. 
We briefly explore the effect of relativistic beaming, gravitational redshift and 
light bending on the angular distribution of disc continuum emission. 
We find that these general relativistic effects do cause the disc to
emit slightly more isotropically, but this is not yet sufficient to produce a
self-consistent model. We discuss wind reprocessing as a potential solution.
Overall, our work suggests that geometric unification
involving an accretion disc wind is a promising scenario, but our results 
pose a number of difficult challenges to such a model.
% Determining the true geometry of ADWs and uncovering the true disc spectral 
% (and angular) energy distribution are key next stpng if we are to build up a 
% holistic picture of the quasar population.
\end{abstract}

@@

The spectra of 
quasars and luminous active galactic nuclei (AGN) 
typically exhibit a series of strong emission lines
with an underlying blue continuum - the so-called {\sl `big blue bump'} (BBB). 
The BBB is normally attributed to emission from a geometrically thin, 
optically thick accretion disc surrounding the central black hole (REF), similar
to that described by \cite{shakurasunyaev1973}.
In addition to the inflowing accreting material, 
outflows are ubiquitous in AGN
and quasars \citep{kellerman1989,ganguly2008}. These outflows can take the form of 
highly collimated radio jets \citep[e.g.][]{hazard1963,potash1980,perley1984,marscher2006}, 
or mass-loaded `winds' emanating from the accretion disc 
\citep{weymann1991,turnermiller2009}. 
Outflows in AGN offer a 
potential feedback mechanism through which the central source can 
affect its environment \citep{king2003,king2005,fabian2012}
-- feedback that is required in models of galaxy evolution (REFs) 
and may explain the `$M-\sigma$' relation \citep{silkrees1998,haring2004}.

Approximately $20\%$ of quasars exhibit blueshifted ($\sim 0.1c$)
broad absorption lines (BALs) in the ultraviolet,
providing clear evidence for outflowing absorbing material
\citep{weymann1991, reichard2003, knigge2008, turnermiller2009, allen2011}.
The simplest explanation for the incidence of 
BAL quasars (BALQSOs) is in terms of an accretion disc wind (ADW). 
Within this paradigm, the BALQSO fraction is associated with
the covering factor of the outflow.
In addition, ADWs may offer a natural explanation for the
diverse phenemonology of luminous AGN and QSOs \citep[e.g.][]{MCGV95, elvis2000}. 
In such a model, a biconical wind rises from 
the accretion disc, and the class of object is explained by the material
intercepting the line of sight. Depending on viewing angle, an observer 
may then see a BALQSO or normal `Type 1' quasar.
Within this framework, the broad-line region (BLR) is normally
assumed to correspond to the dense wind base.
There are various spectroscopic sub-classifications of BALQSOs: 
HiBALQSOs, which only exhibit higher
ionization line absorption; LoBALQSOs which also show
absorption in lower ionization state species such as \mg\ and \al; and
FeLoBALQSOs which show further absorption in Fe~\textsc{ii} and \textsc{iii}.
In unified models, this is generally attributed to ionization stratification
of the outflow \citep[e.g.][]{elvis2000}.

% As well as acting as a source of photoionized plasma, a 
As well as imprinting clear line absorption features,
disc winds may also have a profound effect on the structure and 
emergent {\em continuum} of the accretion disc itself.
Mass-loss will alter the accretion rate and resultant 
temperature of the accretion disc, possibly explaining some 
of the features we typically see in luminous AGN \citep{laordavis2014}.
Recent results from \cite{capellupo2015} find 
that if one includes a combination of mass-loss, general relativity (GR) and Comptonisation
then AGN spectral energy distributions (SEDs) can, in general, be fitted well with accretion disc models.
Mass-loss therefore appears to be critical if an $\alpha$-disc
model is to successfully fit AGN SEDs, particular in the UV region of the spectrum.

@@

The efficiency of line-driving is crucially dependent on the ionization state 
of the outflowing plasma, meaning that it is difficult to prevent 
the wind becoming over-ionized and `failing' in the presence of strong X-rays. 
\cite{MCGV95} proposed a potential solution: 
a region of `hitchhiking gas' that could shield the wind from the central X-ray source. 
Hydrodynamic simulations of line-driven disc winds also found a shielding region
was required to maintain the correct ionization state \citep{PSK2000,PK04}. 
However, \cite{H14} showed that including multiple scattering means the ionizing radiation 
field could still reach the previously shielded regions in those particular models.
An alternative solution is that the wind is clumped \citep[e.g.][]{hamann2013}
possibly on multiple scale lengths. Local density enhancements could lower the 
ionization parameter of the plasma while still maintaining the same mass-loss 
rate and column density.

Evidence for dense substructures in AGN winds is widespread.
BALQSOs show complex absorption line profiles \citep{ganguly2006, simonhamann2010}
and exhibit variability in these profile shapes \citep{capellupo2011,capellupo2012,capellupo2014}.
AGN generally show variability in X-ray absorption components \citep[e.g.][]{risaliti2002}
and many models for the BLR consist of clumps embedded in an outflow 
\citep{krolik1981, emmering1992, dekool1995, cassidyraine1996}.
Clumping can be caused by magnetic confinement \cite{dekool1995},
or the instabilities inherent to line-driven winds 
\citep{lucysolomon1970,macgregor1979,carlberg1980,owockirybicki1984,owockirybicki1985}.
Additionally, clumping is required in line-driven hot star winds 
in order to explain observations \citep{hillier1991eswingsmodel}. Complex substructures 
on a variety of scales are also produced in simulations of line-driven 
outflows in AGN \citep{PSK2000,PK04,progakurosawa2010,proga2014}.
Clumpy winds therefore offer an observationally motivated and theoretically 
predicted way to lower the ionization state of a plasma, possibly in tandem
with a shielding scenario.

@@

There has been some success using simple
kinematic prescriptions for biconical disc winds to model 
AGN and quasar outflows \citep[][hereafter H13]{simlong2008,sim2010,higginbottom2013}. 
H13 successfully produced a benchmark model for BALQSOs. However, the model
had two key drawbacks. Firstly, an unrealistically low X-ray luminosity
was required in order to prevent over-ionization of the outflow.
Secondly, the model failed to produce the strong emission lines 
required at low inclinations in a unified model.
In this paper, we attempt to address these issues, and test the disc wind 
unification model using Monte Carlo radiative transfer (MCRT) and photoionization 
calculations. The paper is organised as follows.
In section 2, we describe some of the important photoionization 
and MCRT aspects of the code. In section 3, we outline the model, including 
a description of our clumping implementation. In section 4, we present the results 
from a clumped model. In section 5 we discuss our results, 
focussing particular on the anisotropy of 
disc emission and GR effects, and finally, in section 6, we summarise our findings.

@@

We use the MCRT code \py\ to carry out our radiative transfer and photoionization
calculations in non-local thermodynamic equilibrium (non-LTE). 
The has been used to model accreting white dwarfs 
(Long \& Knigge 2002, hereafter LK02; Noebauer et al. 2010; 
Matthews et al. 2015, hereafter M15), young-stellar objects \citep{simmacro2005}
and quasars (H13). \nocite{noebauer, M15, LK02}
These studies contain extensive detail on the code, 
so we only briefly describe the key elements of the global 
ionization calculation and other important aspects.

@@

%To avoid approximating the complex AGN SED as e.g. a dilute blackbody \citep[][M15]{ML93},
We model the SED in a cell using the technique described by H13. In this scheme,
the mean intensity in a series of bands is modeled as a normalised power law in 
frequency $\nu$
\begin{equation}
J_{\nu,j}=K_{pl}\nu^{\alpha_{pl}},
\end{equation}
for a band $j$, or an exponential 
\begin{equation}
J_{\nu,j}=K_{exp} e^{(-h\nu/k T_{exp})}.
\end{equation}
Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
deduced from the band-limited radiation field estimators.
The ionization rate out of ion $i$ can then be written as 
\begin{equation}
R_{i,i+1}(J)= 
\displaystyle{
n_i \left(C_{i} n_e + 
\sum_{band~j=0}^{n}~
{\int_{\nu_j}^{~\nu_j+1}{J_{\nu,j}\sigma_i(\nu)\nu^{-1}d\nu}}
\right),}
\end{equation}
where $\sigma_i$ is the photoionization cross-section and $C_{i}$
represents the collisional ionization coefficient.
The recombination rate {\em into} ion $i$ is given by 
\begin{equation}
R_{i+1,i}(T_e) = (\alpha^i_{RR} + \alpha^i_{DR} + \alpha^i_{CR}) n_{i+1} n_e,
\end{equation}
where each $\alpha^i$ here is the recombination rate coefficient into the ground state of ion $i$.
The subscripts denote radiative, dielectronic and 
collisional (three-body) recombination, respectively.
We neglect recombination to and from excited states in the simple-ion calculation. 
For simple-ions, we use a dilute Boltzmann equation to calculate 
the population of level $k$ in ionic stage $j$,
\begin{equation}
\frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
\end{equation}
Here $z_j$ is the partition function of ionic stage $j$,
$T_R$ is the effective radiation temperature,
$E_k$ is the energy difference between level $k$ and the ground state,
and $g_k$ is the statistical weight of level $k$.

@@

The accretion disc in our model is geometrically thin, but optically thick.
We thus adopt a standard multi-temperature blackbody
using a \cite{shakurasunyaev1973} temperature profile.

The emergent SED is then determined by the specified accretion rate ($\dot{m}$)
and central BH mass ($M_{BH}$).
All photon sources in our model are assumed to beopaque, meaning
that photons which strike them are destroyed.
The inner radius of the disc extends to the innermost 
stable circular orbit (ISCO) of the BH. 
We assume a Schwarzchild BH with an ISCO at $6~R_G$.
The X-ray source in our model is treated as an isotropic sphere at the ISCO,
which emits r-packets according to a power law in flux with index $\alpha_X$, of the form
\begin{equation}
F_X (\nu) = K \nu^{\alpha_X}.
\end{equation}
The normalisation, $K_X$ of this power law is such that it 
produces the specified 2-10~keV luminosity, $L_X$.
In addition to the disc and X-ray source, 
the wind is able to reprocess radiation. However, new 
photon packets are not produced in the wind (as in LK02). 
Instead, this reprocessing is dealt with by enforcing strict
radiative equilibrium ({\em modulo} adiabatic cooling)
via an indivisible energy packet
constraint (see Lucy 2002, M15; also section 2.3).
% \begin{equation}

@@

To allow for clumping in our outflow we adopt a simple approximation
used extensively in stellar wind modelling, known as 
{\em microclumping} \citep{hamann1998}. The key assumption here is that 
typical clump sizes are much smaller than the typical photon mean free path, 
and thus the clumps are 
both geometrically and optically thin. This approach 
allows one to introduce a `filling factor', $f$, which is the 
fraction of the volume of the plasma filled by clumps. The intra-clump
medium is assumed to be a vacuum. We can then introduce the 
density enhancement, $D$, which is defined as 
\begin{equation}
D = \frac{1}{f}.
\end{equation}
We then multiply all densities in the model by $D$, and all emitting volumes
by $f$, meaning that all $\rho^2$ emissivities and opacities
(such as collisional excitation and recombination) will be enhanced, 
while all $\rho$ process (such as electron scattering and bound-free opacity)
remain unchanged for a given ionization state. 
% A factor $f$ is also
% applied to the opacities such that opacities which scale only with $\rho$ are not
% increased.

@@

\section{Results}

Here we describe the results from our model, the parameters
of which are shown in table~1.
Parameters differing from the benchmark model of
H13 are highlighted with an asterisk.
This set of parameters
was arrived at by conducting a limited grid search over a 
5-dimensional parameter space involving the variables
$r_{min}$, $\theta_{min}$, $f$, $\alpha$ and $R_V$.
The full grid, including output spectral files and plots can be found at
\url{jhmatthews.github.io/quasar-wind-grid/}.
We evaluated these models qualitatively based on the following
criteria:
\begin{itemize}
\item Does the model maintain the correct ionization state and produce strong BALs?
\item Does strong line emission emerge at low inclinations?
\item Do H recombination lines appear in the spectrum?
\item Do a certain range of angles produce LoBAL features?
\item Does the model compare favourably to quasar composite spectra?
\end{itemize}
We arrived at a model which is one of the most promising,
but also broadly representative of a family of models. 
Examination of this individual model allows us to gain insight
into fundamental geometrical and physical constraints, 
as discussed in section 5, and assess the models potential
for unification.

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/wind.png}
\caption
{
Physical properties of the outflow, shown by the coloured contours.
The solid black line marks a sphere at $1000~R_G$.
The dotted lines show the $72^\circ$ and $78^\circ$ sightlines 
to the centre of the system, and illustrate that different sightlines
intersect material of different ionization states.
}
\label{fig:wind}
\end{figure*} %fullpage

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/sightlines.png}
\caption
{
The classes of sightlines and the corresponding possible 
system types in the framework of our model.
%[FIGURE NEEDS IMPROVING]
}
\label{fig:sightline}
\end{figure}

@@

One commonly used measure of the ionization state is the ionization parameter, $U_H$, given by
\begin{equation}
U_H = \frac{4\pi}{n_H c}\int_{13.6{\rm{eV}}}^{\infty}\frac{{J_{\nu}d\nu}}{h\nu}.
\end{equation}
\noindent where $\rm{n_H}$ is the local number density of H, and $\nu$ denotes photon 
frequency. The ionization parameter represents the ratio of the number density of 
H ionizing photons to the local H density. It is, however, a poor measure of the 
ionization state of species such as \civ\ as it encodes no information
about the shape of the SED. In our case, the X-ray photons 
are dominant in the photoionization of the UV resonance line ions. 
This explains why a factor of 100 increase in X-ray luminosity requires
a clumping factor of 0.01, even though the value of $U_H$ decreases by only a factor of $\sim10$ 
compared to H13. This is also the reason for significant Lyman edge photoabsorption
at the highest inclinations (see section 4.3).

Clumping also causes the total line luminosity to increase dramatically,
as recombination and collisional excitation are both proportional to
$n_e^2$. This line emission typically emerges on the edge of the wind
nearest the central source. The location of the line emitting regions
is dependent on the ionization state, as well as the X-rays heating the plasma.
The radii of these emitting regions is important,
and can be compared to observations. As shown in figure~\ref{fig:wind},
the \civ\ line in our model is typically formed between 
$100-1000~R_G$ ($\sim10^{17}-10^{18}~\rm{cm}$).
This is in rough agreement with the reverberation mapping 
results of Kaspi (2000) for the $2.6\times10^{9} M_\odot$ quasar S5 0836+71,
and also compares favourably with microlensing measurements of the size of the
\civ\ emission line region in the BALQSO H1413+117 \citep{odowd2015}.

@@

@@

\section{Discussion: The Disc SED}

>>>>

\begin{abstract} 
% Broad absorption lines (BALs) in the ultraviolet 
% are seen in $\sim20\%$ of quasi-stellar objects (QSOs). 
% Blue-shifted broad absorption lines (BALs) are the most direct evidence of 
% accretion disc `winds' in such systems; mass loaded outflows
% emanating from the disc that may be driven by line forces or
% magnetic processes. 
% Various unification schemes for
% quasars and luminous active galactic nuclei (AGN) have proposed
% that the broad emission line region is roughly cospatial
% with broad absorption line (BAL) gas and much of the phenomenology of luminous AGN
% can be explained by a simple geometrical picture involving an accretion
% disc and associated outflow. Here, we test this paradigm by 
% utilising our state-of-the-art radiative transfer code to produce synthetic spectra
% from simple biconical disc wind models. 
% In particular, we expand on our previous 
% work in which a benchmark model for BAL quasars was produced. 
% We have conducted a limited parameter search with the aim of unifying quasar phenomenology,
% and arrive at an improved model. The grid is now publicly available.
% We find that a simple treatment of clumping (`microclumping') 
% allows for a more realistic X-ray luminosity in the model by lowering the 
% ionization parameter. We examine the X-ray properties of this new model
% and find good agreement with existing X-ray samples of AGN and QSOs.
% We find that the dense, X-ray heated wind 
% produces strong H recombination and collisionally excited resonance 
% line emission to emerge at the low inclination angles, 
% which represent quasars within this unification scenario.
% % We also treat Hydrogen using a `macro-atom' approach in order to 
% % examine the effect of recombination on Hydrogen emission lines, and
% % this results in significant line emission in Lyman~$\alpha$ and the Balmer series.
% However, we are unable to reproduce the remarkably similar of line-to-continuum
% ratios between BAL and non-BAL quasar composites.
% This is due to a fundamental constraint arising
% from the anisotropy of emission from a classical thin disc. 
% We briefly explore the effect of relativistic beaming, gravitational redshift and 
% light bending on the angular distribution of disc continuum emission. 
% We find that these general relativistic effects do cause the disc to
% emit slightly more isotropically, but this is not yet sufficient to produce a
% self-consistent model. We discuss wind reprocessing as a potential solution.
% Overall, our work suggests that geometric unification
% involving an accretion disc wind is a promising scenario, but our results 
% pose a number of difficult challenges to such a model.
% Determining the true geometry of ADWs and uncovering the true disc spectral 
% (and angular) energy distribution are key next stpng if we are to build up a 
% holistic picture of the quasar population.
Various unification schemes for quasars and luminous active galactic nuclei (AGN) 
propose that much of their complex phenomenology can be explained by a simple 
geometrical picture involving an accretion disc and associated outflow.
Here, we test this paradigm by utilising our state-of-the-art radiative transfer 
and photoionization code to produce synthetic spectra from simple biconical disc wind models. 
In particular, we expand on our previous work by presenting a model that
exhibits many of the spectral features expected from a quasar unification model.
We find that a simple treatment of clumping (`microclumping') 
allows for a more realistic X-ray luminosity in the model,
while maintaining the ionization state necessary for strong BAL features. 
We examine the X-ray properties of this new model
and find good agreement with existing X-ray samples of AGN and quasars.
We find that the dense, X-ray heated wind 
produces strong recombination and collisionally excited line emission in, 
e.g., \civ\ and \la, to emerge at the low inclination, `Type 1 quasar-like' angles. 
The highest inclination models possess prominent \mg\ and \al\ BALs, the absorption 
features seen in LoBAL quasars. Despite these successes, we are unable to reproduce 
the remarkably uniform emission line properties 
% remarkably similar of line-to-continuum ratios apparent 
seen in BAL and non-BAL quasar composites.
This is due to a fundamental constraint arising
from the anisotropy of emission from a classical thin disc. 
We briefly explore the impact of general relativistic effects
on the angular distribution of disc continuum emission, and find that 
they are not sufficient to produce a self-consistent model.
% relativistic beaming, gravitational redshift and 
% light bending on the angular distribution of disc continuum emission. 
% We find that these general relativistic effects are not yet sufficient to produce a
% self-consistent model. 
Overall, our work suggests that geometric unification
involving an accretion disc wind is a promising scenario, but our results 
pose a number of difficult challenges to such a model.
\end{abstract}

@@

The spectra of 
quasars and luminous active galactic nuclei (AGN) 
typically exhibit a series of strong emission lines
with an underlying blue continuum - the so-called {\sl `big blue bump'} (BBB). 
The BBB is often attributed to emssion from a 
geometrically thin, optically thick accretion disc surrounding the central black hole, similar
to that described by \cite{shakurasunyaev1973}.
In addition to the inflowing accreting material, i
outflows are ubiquitous in AGN
and quasars \citep{kellerman1989,ganguly2008}. These outflows can take the form of 
highly collimated radio jets \citep[e.g.][]{hazard1963,potash1980,perley1984,marscher2006}, 
or mass-loaded `winds' emanating from the accretion disc 
\citep{weymann1991,turnermiller2009}. 
Outflows in AGN offer a 
potential feedback mechanism through which the central source can 
affect its environment \citep{king2003,king2005,fabian2012}
-- feedback that is required in models of galaxy evolution \citep{springel2005}
and may explain the `$M-\sigma$' relation \citep{silkrees1998,haring2004}.

Approximately $20\%$ of quasars exhibit blueshifted ($\sim 0.1c$)
broad absorption lines (BALs) in the ultraviolet,
providing clear evidence for outflowing absorbing material
\citep{weymann1991, reichard2003, knigge2008, turnermiller2009, allen2011}.
The simplest explanation for the incidence of 
BAL quasars (BALQSOs) is in terms of an accretion disc wind (ADW). 
Within the ADW paradigm, a biconical wind rises from 
the accretion disc and the BALQSO fraction is associated with
the covering factor of the outflow. 
The wind is expected to be roughly equatorial from polarisation results
\citep{goodrich1995, cohen1995}.
ADWs offer a natural explanation for the
diverse phenemonology of luminous AGN and QSOs \citep[e.g.][]{MCGV95, elvis2000}. 
Depending on viewing angle, an observer 
may see a BALQSO or normal `Type 1' quasar.
Within this framework, the broad-line region (BLR) can 
correspond either to the dense wind base or dense clumps embedded
in the outflows. Indeed, \cite{elitzur2014} found that a disc-wind BLR scenario
naturally explains the emission line evolution of AGN.
A biconical wind model can also readily explain the various sub-classifications of BALQSOs: 
HiBALQSOs, which only exhibit higher ionization line absorption; LoBALQSOs which also show
absorption in lower ionization state species such as \mg\ and \al; and
FeLoBALQSOs which show further absorption in Fe~\textsc{ii} and \textsc{iii}.
In unified models, this is generally attributed to ionization stratification
of the outflow \citep[e.g.][]{elvis2000}.

% As well as acting as a source of photoionized plasma, a 
As well as imprinting clear line absorption and emission
features, disc winds may also have a profound effect on the structure and 
emergent {\em continuum} of the accretion disc itself.
Mass-loss will alter the accretion rate and resultant 
temperature of the accretion disc, possibly explaining some 
of the features typically seen in luminous AGN \citep{knigge1999,laordavis2014}.
There have been numerous difficulties when confronting 
theoretical accretion disc models with observations 
\citep[see e.g.][]{blaes1998}. However, AGN spectral energy distributions (SEDs) can now, 
in general, be fitted well with accretion disc models
when the effects of general relativity (GR), Comptonisation
and mass-loss are included \citep{capellupo2015}. Mass-loss therefore appears to be 
critical if an accretion disc model is to successfully fit AGN SEDs, 
particular in the UV region of the spectrum.

@@

The efficiency of line-driving is crucially dependent on the ionization state 
of the outflowing plasma, meaning that it is difficult to prevent 
the wind becoming over-ionized and `failing' in the presence of strong X-rays. 
\cite{MCGV95} proposed a potential solution: 
a region of `hitchhiking gas' that could shield the wind from the central X-ray source. 
Hydrodynamic simulations of line-driven disc winds also found a shielding region
was required to maintain the correct ionization state \citep{PSK2000,PK04}. 
However, \cite{H14} showed that including multiple scattering means the ionizing radiation 
field could still reach the previously shielded regions in those particular models.
An additional or alternative solution is that the wind is clumped \citep[e.g.][]{hamann2013}
possibly on multiple scale lengths. Local density enhancements could lower the 
ionization parameter of the plasma while still maintaining the same mass-loss 
rate and column density.

Evidence for dense substructures in AGN winds is widespread.
BALQSOs show complex absorption line profiles \citep{ganguly2006, simonhamann2010}
and exhibit variability in these profile shapes \citep{capellupo2011,capellupo2012,capellupo2014}.
AGN generally show variability in X-ray absorption components \citep[e.g.][]{risaliti2002}
and many models for the BLR consist of clumps embedded in an outflow 
\citep{krolik1981, emmering1992, dekool1995, cassidyraine1996}.
Clumping can be caused by magnetic confinement \cite{dekool1995},
or the instabilities inherent to line-driven winds 
\citep{lucysolomon1970,macgregor1979,carlberg1980,owockirybicki1984,owockirybicki1985}.
Additionally, clumping is required to explain the electron scattering wings of emission lines formed
in line-driven hot star winds \citep{hillier1991eswingsmodel}. Complex substructures 
on a variety of scales are also produced in simulations of line-driven 
outflows in AGN \citep{PSK2000,PK04,progakurosawa2010,proga2014}.
Clumpy winds therefore offer an observationally motivated and theoretically 
predicted way to lower the ionization state of a plasma, possibly in tandem
with a shielding scenario.

@@

There has been some success using simple
kinematic prescriptions for biconical disc winds to model 
AGN and quasar outflows \citep[][hereafter H13]{simlong2008,sim2010,higginbottom2013}. 
H13 successfully produced a benchmark model for BALQSOs. However, the model
had two key drawbacks. First, an unrealistically low X-ray luminosity
was required in order to prevent over-ionization of the outflow.
Second, the model failed to produce the strong emission lines 
required at low inclinations in a unified model.
In this paper, we attempt to address these issues, and test the disc wind 
unification model using Monte Carlo radiative transfer (MCRT) and photoionization 
calculations. The paper is organised as follows.
In section 2, we describe some of the important photoionization 
and MCRT aspects of the code. In section 3, we outline the model, including 
a description of our clumping implementation. In section 4, we present the results 
from a clumped model. In section 5 we discuss our results, 
focussing particular on the anisotropy of 
disc emission and GR effects, and finally, in section 6, we summarise our findings.

@@

We use the MCRT code \py\ to carry out our radiative transfer and photoionization
calculations in non-local thermodynamic equilibrium (non-LTE). 
The code has been used to model accreting white dwarfs 
(Long \& Knigge 2002, hereafter LK02; Noebauer et al. 2010; 
Matthews et al. 2015, hereafter M15), young-stellar objects \citep{simmacro2005}
and quasars (H13). \nocite{noebauer, M15, LK02}
These studies contain extensive detail on the code, 
so we only briefly describe the key elements of the global 
ionization calculation and other important aspects.

@@

%To avoid approximating the complex AGN SED as e.g. a dilute blackbody \citep[][M15]{ML93},
We model the SED in a cell using the technique described by H13. In this scheme,
the mean intensity in a series of $n$ bands is modeled as a normalised power law in 
frequency $\nu$
\begin{equation}
J_{\nu,i}=K_{pl}\nu^{\alpha_{pl}},
\end{equation}
for a band $i$, or an exponential 
\begin{equation}
J_{\nu,i}=K_{exp} e^{(-h\nu/k T_{exp})}.
\end{equation}
Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
deduced from the band-limited radiation field estimators.
The ionization rate out of ion $j$ can then be written as 
\begin{equation}
R_{j,j+1}(J)= 
\displaystyle{
n_i \left(C_{j} n_e + 
\sum_{band~i=0}^{n}~
{\int_{\nu_i}^{~\nu_i+1}{J_{\nu,i}\sigma_j(\nu)\nu^{-1}d\nu}}
\right),}
\end{equation}
where $\sigma_j$ is the photoionization cross-section and $C_{j}$
represents the collisional ionization coefficient.
The recombination rate {\em into} ion $j$ is given by 
\begin{equation}
R_{j+1,j}(T_e) = (\alpha^j_{RR} + \alpha^j_{DR} + \alpha^j_{CR}) n_{j+1} n_e,
\end{equation}
where each $\alpha^j$ here is the recombination rate coefficient into the ground state of ion $j$.
The subscripts denote radiative, dielectronic and 
collisional (three-body) recombination, respectively.
We neglect recombination to and from excited states in the simple-ion calculation. 
For simple-ions, we use a dilute Boltzmann equation to calculate 
the population of level $k$ in ionic stage $j$,
\begin{equation}
\frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
\end{equation}
Here $z_j$ is the partition function of ionic stage $j$,
$T_R$ is the effective radiation temperature,
$E_k$ is the energy difference between level $k$ and the ground state,
and $g_k$ is the statistical weight of level $k$.

@@

The accretion disc in our model is geometrically thin, but optically thick.
We thus treat the disc as an ensemble of blackbodies with a 
\cite{shakurasunyaev1973} effective temperature profile. 
The emergent SED is then determined by the specified accretion rate ($\dot{m}$)
and central BH mass ($M_{BH}$).
All photon sources in our model are assumed to beopaque, meaning
that photons which strike them are destroyed.
The inner radius of the disc extends to the innermost 
stable circular orbit (ISCO) of the BH. 
We assume a Schwarzchild BH with an ISCO at $6~r_G$, where 
$r_G = GM_{BH}/c^2$ is the gravitational radius.
For a $10^9~M_\odot$ black hole, this is equal to $8.8\times10^{14}~{\rm cm}$ 
or $\sim10^{-4}~{\rm pc}$.

The X-ray source in our model is treated as an isotropic sphere at the ISCO,
which emits r-packets according to a power law in flux with index $\alpha_X$, of the form
\begin{equation}
F_X (\nu) = K \nu^{\alpha_X}.
\end{equation}
The normalisation, $K_X$ of this power law is such that it 
produces the specified 2-10~keV luminosity, $L_X$.
In addition to the disc and X-ray source, 
the wind is able to reprocess radiation. However, new 
photon packets are not produced in the wind (as in LK02). 
Instead, this reprocessing is dealt with by enforcing strict
radiative equilibrium ({\em modulo} adiabatic cooling; see section~2.3)
via an indivisible energy packet
constraint (see Lucy 2002, M15).
% \begin{equation}

@@

To allow for clumping in our outflow we adopt a simple approximation
used extensively in stellar wind modelling, known as 
{\em microclumping} \citep{hamann1998}. The key assumption here is that 
typical clump sizes are much smaller than the typical photon mean free path, 
and thus the clumps are 
both geometrically and optically thin. This approach 
allows one to introduce a `filling factor', $f$, which is the 
fraction of the volume of the plasma filled by clumps. The intra-clump
medium is assumed to be a vacuum. We can then introduce the 
density enhancement, $D$, which is defined as 
\begin{equation}
D = \frac{1}{f}.
\end{equation}
We then multiply all densities in the model by $D$, and all emitting volumes
by $f$. This has the effect of enhancing all emissivities and opacities
that scale with the square of density (such as collisional excitation and recombination) 
by a factor $D$. All processes that scale linearly with density 
(such as electron scattering and bound-free opacity)
will remain unchanged for a given ionization state. 
% A factor $f$ is also
% applied to the opacities such that opacities which scale only with $\rho$ are not
% increased.

@@

\section{Results and Discussion}

Here we describe the results from our model, the parameters
of which are shown in table~1.
Parameters differing from the benchmark model of
H13 are highlighted with an asterisk.
This set of parameters
was arrived at by conducting a limited grid search over a 
5-dimensional parameter space involving the variables
$r_{min}$, $\theta_{min}$, $f$, $\alpha$ and $R_V$.
The full grid, including output spectral files and plots can be found at
\url{jhmatthews.github.io/quasar-wind-grid/}.
We evaluated these models qualitatively based on the following
criteria:
\begin{itemize}
\item Does the model maintain the right ionization state to produce strong BALs?
\item Does significant line emission emerge at low inclinations?
\item Do H recombination lines appear in the spectrum?
\item Do a certain range of angles produce LoBAL features?
\item Does the model compare favourably to quasar composite spectra?
\end{itemize}
In this section, we present one of the most promising models and discuss
the various successes and failures.
This allows us to gain insight into fundamental geometrical 
and physical constraints and assess the potential for unification.

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/link.png}
\caption
{
Physical properties of the outflow, shown by the coloured contours.
The solid black line marks a sphere at $1000~r_G$.
The dotted lines show the $72^\circ$ and $78^\circ$ sightlines 
to the centre of the system, and illustrate that different sightlines
intersect material of different ionization states.
}
\label{fig:wind}
\end{figure*} %fullpage

@@

One common way to quantify the ionization state of a plasma
is through the ionization parameter, $U_H$, given by
\begin{equation}
U_H = \frac{4\pi}{n_H c}\int_{13.6{\rm{eV}}}^{\infty}\frac{{J_{\nu}d\nu}}{h\nu}.
\end{equation}
\noindent where $\rm{n_H}$ is the local number density of H, and $\nu$ denotes photon 
frequency. Shown in figure~\ref{fig:wind},
the ionization parameter is a useful measure of the global ionization state,
as it represents the ratio of the number density of 
H ionizing photons to the local H density.
It is, however, a poor representation of the 
ionization state of species such as \civ\ as it encodes no information
about the shape of the SED. In our case, the X-ray photons 
are dominant in the photoionization of the UV resonance line ions. 
This explains why a factor of 100 increase in X-ray luminosity requires
a clumping factor of 0.01, even though the value of $U_H$ decreases by only a factor of $\sim10$ 
compared to H13. This is also the reason for significant Lyman edge photoabsorption
at the highest inclinations (see section 4.3).

Clumping also causes the total line luminosity to increase dramatically,
as recombination and collisional excitation are both proportional to
$n_e^2$. This line emission typically emerges on the edge of the wind
nearest the central source. The location of the line emitting regions
is dependent on the ionization state, as well as the X-rays heating the plasma.
The radii of these emitting regions is important,
and can be compared to observations. As shown in figure~\ref{fig:wind},
the \civ\ line in our model is typically formed between 
$100-1000~r_G$ ($\sim10^{17}-10^{18}~\rm{cm}$).
This is in rough agreement with the reverberation mapping 
results of Kaspi (2000) for the $2.6\times10^{9} M_\odot$ quasar S5 0836+71,
and also compares favourably with microlensing measurements of the size of the
\civ\ emission line region in the BALQSO H1413+117 \citep{odowd2015}.

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/windnew3.png}
\caption
{
A cartoon describing the broad classes of sightline 
in our model, illustrates how geometric effects lead to 
the different emergent spectra. The colour gradient is approximate,
but indicates the stratified ionization structure, 
from highly ionized (yellow) to low ionization (purple) material.
%[FIGURE NEEDS IMPROVING]
}
\label{fig:sightline}
\end{figure}

@@

% This could potentially be achieved by wind reprocessing or a more realistic, GR treatment
% of the accretion disc. We briefly discuss both these possibilities.

% \subsection{Wind reprocessing}
